---
title: "Game of Life"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Game of Life}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

[Conway's Game of Life](https://en.wikipedia.org/wiki/Conway%27s_Game_of_Life)
is a cellular automaton where the evolution of the grid state is determined
entirely by the initially state of the grid.

If we consider the grid to be black pixels on a white background, at each
time step the boart updates by following these rules:

1. Any live cell with two or three live neighbours survives.
2. Any dead cell with three live neighbours becomes a live cell.
3. All other live cells die in the next generation. Similarly, all other dead cells stay dead.

## General implementation notes for R

A naive approach to implemention of this game would be to have nested `for`
loops and iterate over every cell.  At every cell then, interrogate all the
neighbours to decide on what the cell should do for the next time step.

Nested `for` loops in R are too slow to produce a smoothly animating result.
Instead, a matrix-based approach is taken where the board state is
shifted in the 8 directions, and then these shifted boards are stacked
and summed efficiently to determine the next board state.

## Rendering in the Event Loop

At every frame, the board is cleared and the entire board state is drawn
from scratch using `grid::grid.raster()` to plot a matrix to the graphics
output device.

Also at every frame, a new board state is calculated.  If a `target_fps` is
not set, and the simulation allowed to run as fast as possible, then the 
board is updated over 100 times per second.

## Interactivity

This application watches for a mouse click by the user.  When the mouse
is held down, the board updates are paused, and the pixels nearest the 
mouse position are set to "active".  


# Game of Life in R

```{r setup, eval = FALSE}
library(eventloop)
library(grid)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Global State
#  - 'N'     size of board
#  - 'g'     the game board is an NxN integer matrix of 0/1 values
#  - 'latch' is set to TRUE whenever the user presses-and-holds a mouse button
#            and then set to FALSE when the mouse button is released
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
N <- 100
g <- matrix(sample(c(0L, 1L), N*N, replace = TRUE), N, N)
latch <- FALSE


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Standard blank vectors needed when shifting the current board in 8 
# different directions
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
blankN   <- rep(0L, N   )
blankNm1 <- rep(0L, N-1L)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' The gave of life update function.
#' 
#' Press ESC to quit.
#'
#' @param event The event from the graphics device. Is NULL when no event
#'        occurred.  Otherwise has `type` element set to:
#'        `event$type = 'mouse_down'` 
#'               - an event in which a mouse button was pressed
#'               - `event$button` gives the index of the button
#'        `event$type = 'mouse_up'`   
#'               - a mouse button was released
#'        `event$type = 'key_press'`  
#'               - a key was pressed
#'               - `event$char` holds the character as string
#'               - `event$int` holds the integer representation
#' @param mouse_x,mouse_y current location of mouse within window. If mouse is 
#'        not within window, this will be set to the last available coordinates
#' @param frame_num integer count of which frame this is
#' @param fps_actual,fps_target the curent framerate and the framerate specified
#'        by the user
#' @param dev_width,dev_height the width and height of the output device. Note:
#'        this does not cope well if you resize the window
#' @param ... any extra arguments ignored
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
game_of_life <- function(event, mouse_x, mouse_y, ...) {

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Set the 'latch' variable depending on whether or not a button is pressed
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (!is.null(event)) {
    if (event$type == 'mouse_down') {
      latch <<- TRUE
    } else if (event$type == 'mouse_up') {
      latch <<- FALSE
    }
  }

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # if 'latch' is TRUE, then set the pixel under the mouse to 'black',
  # Otherwise update the board 
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (latch) {
    g[as.integer((1 - mouse_y) * (N-1) + 1), as.integer(mouse_x * (N-1) + 1)] <<- 1L
  } else {
    
    # Neighbours, everybody needs good neighbours....
    #  1 2 3
    #  4   5
    #  6 7 8
    gsum <- cbind(blankN, rbind(blankNm1, g[-N, -N])) +  # 1
      rbind(blankN, g[-N,]) +                            # 2
      cbind(rbind(blankNm1, g[-N, -1]), blankN) +        # 3
      cbind(blankN, g[,-N]) +                            # 4
      cbind(g[,-1], blankN) +                            # 5
      cbind(blankN, rbind(g[-1, -N], blankNm1)) +        # 6
      rbind(g[-1, ], blankN) +                           # 7
      cbind(rbind(g[-1, -1], blankNm1), blankN)          # 8
    
    # Standard Conway rules
    g[] <<- (g == 1 & (gsum == 2L | gsum == 3)) | (g == 0 & gsum == 3)
    
  }
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Clear the screen (Note: grid.rect() is faster than a grid.newpage())
  # Draw the GameOfLife grid
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  grid::grid.rect(gp = grid::gpar(fill = 'black'))
  grid::grid.raster(1L - g, interpolate = FALSE)
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run the loop
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
run_loop(game_of_life, 7, 7, fps_target = 30, show_fps = TRUE)
```


Since an interactive window cannot be captured in a vignette, a video
screen capture has been taken of the window and included below.

<video controls>
  <source src="images/game-of-life.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 
