---
title: "Spirograph"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spirograph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## A Four-fold symmetric spirograph drawing application

* Left mouse is used to active drawing mode
* Pressing SPACE will clear the canvas
* Adjust drawing mode with PLUS and MINUS for size, 0-9 for colour and 
  'm' for drawing object
* extra logic is used to store prior coordinates in the global state so 
  that the output is a continuous line when drawing
* drawing elements are then repeated in 4 quadrants

## Event Loop Considerations



```{r setup, eval=FALSE}
library(grid)
library(eventloop)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Set up the global variables which store the state of the world
#  'drawing'  Currently drawing?
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
drawing  <- FALSE
mode     <- 1L
mode_max <- 1L
last_x   <- NA
last_y   <- NA
size     <- 2
colours  <- c('black', 'darkgreen', 'darkblue', 'lightblue', 'brown2', 
              'red', 'hotpink', 'darkgray', 'gold', 'white')
colour   <- colours[[1]]

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# The main 'draw' function - his function is called repeatedly within the eventloop.
#
# If 'event' is not NULL, then it means that the user interacted with the
# display.  The following events have an effect on the canvas:
#  - hold mouse to set drawing mode
#  - releasing the mouse button stops drawing mode
#  - pressing SPACE clears the canvas
#  - +/- changs the size of the object drawn
#  - 'm' changes the type ofelement drawn
#  - 0-9 changes the drawing colour
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
draw <- function(event, mouse_x, mouse_y, ...) {
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Process events
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (!is.null(event)) {
    if (event$type == 'mouse_down') {
      drawing <<- TRUE
    } else if (event$type == 'mouse_up') {
      drawing <<- FALSE
      last_x <<- NA
      last_y <<- NA
    } else if (event$type == 'key_press') {
      if (event$char == ' ') {
        grid::grid.rect(gp = gpar(col=NA, fill='white')) # clear screen
        last_x <<- NA
        last_y <<- NA
      } else if (event$char == 'm') {
        mode <<- (mode + 1L) %% (mode_max + 1L)
        cat("Mode: ", mode, "\n")
      } else if (event$char == '+') {
        size <<- size + 2
        cat("Size: ", size, "\n")
      } else if (event$char == '-') {
        size <<- size - 2
        if (size < 2) {
          size <<- 2
        }
        cat("Size: ", size, "\n")
      } else if (event$char %in% 0:9) {
        colour <<- colours[[as.integer(event$char) + 1L]]
        cat("Colour: ", colour, "\n")
      }
    }
  }
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # If drawing mode is active, draw from the last point to this point.
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (drawing) {
    if (!is.na(last_x)){
      x <- c(last_x, mouse_x)
      y <- c(last_y, mouse_y)
      
      mx = c(x, 1-x, 1-x,   x)
      my = c(y,   y, 1-y, 1-y)
      
      if (mode == 0) {
        grid::grid.polyline(
          x = c(x, 1-x, 1-x,   x),
          y = c(y,   y, 1-y, 1-y),
          id.lengths = c(2, 2, 2, 2),
          gp = gpar(
            lwd = size,
            col = colour
          )
        )
      } else if (mode == 1) {
        grid::grid.circle(
          x = mx,
          y = my,
          r = unit(size, 'pt'),
          gp = gpar(col = NA, fill = colour)
        )
      }
    }
    
    last_x <<- mouse_x
    last_y <<- mouse_y
  }
  
  
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Start the event loop.  
# Press ESC to quit
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
eventloop::run_loop(draw, double_buffer = FALSE)
```


<video controls>
  <source src="images/spirograph.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 

