---
title: "Asteroids"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Asteroids}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## The 'Asteroids' arcade game

Control a ship and avoid the asteroids!

Note: No collision mechanism coded yet - so objects just pass through each other.

## Controls

* LEFT and RIGHT to rotate the ship
* UP and DOWN to accelerate/decelerate the ship.
* Press ESC to quit




```{r setup, eval=FALSE}
library(eventloop)
library(grid)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Ship
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
ship_poly_x <- unit(c(- 7,   7,  0), 'pt')
ship_poly_y <- unit(c(-10, -10, 15), 'pt')

angle <- 0
xpos  <- 0.5
ypos  <- 0.5
xvel  <- 0
yvel  <- 0

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Asteroid
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
asteroid_x <- unit(c(1, 1, 4, 3, 5, 9, 9, 7, 10, 8, 7, 4) - 5.6, 'pt') * 10
asteroid_y <- unit(c(5, 7, 7, 9, 9, 8, 7, 5,  3, 1, 2, 1) - 5.3, 'pt') * 10

Nasteroids <- 4; set.seed(1)
asxpos <- runif(Nasteroids)
asypos <- runif(Nasteroids)
asxvel <- runif(Nasteroids)/200
asyvel <- runif(Nasteroids)/200


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Update the positions of the asteroids and the users ship in the eventloop
#' 
#' @param event The event from the graphics device. Is NULL when no event
#'        occurred.  Otherwise has `type` element set to:
#'        `event$type = 'mouse_down'` 
#'               - an event in which a mouse button was pressed
#'               - `event$button` gives the index of the button
#'        `event$type = 'mouse_up'`   
#'               - a mouse button was released
#'        `event$type = 'mouse_move'`   
#'               - mouse was moved 
#'        `event$type = 'key_press'`  
#'               - a key was pressed
#'               - `event$char` holds the character as string
#'               - `event$int` holds the integer representation
#' @param mouse_x,mouse_y current location of mouse within window. If mouse is 
#'        not within window, this will be set to the last available coordinates
#' @param frame_num integer count of which frame this is
#' @param fps_actual,fps_target the curent framerate and the framerate specified
#'        by the user
#' @param dev_width,dev_height the width and height of the output device. Note:
#'        this does not cope well if you resize the window
#' @param ... any extra arguments ignored
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
asteroids <- function(event, mouse_x, mouse_y, frame_num, fps_actual,
                         fps_target, dev_width, dev_height, ...) {
  grid::grid.rect(gp = grid::gpar(fill = 'black'))

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Keyboard control Left/Right/Up
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (!is.null(event) && event$type == 'key_press') {
    if (event$char == 'Left') {
      angle <<- angle + 8 * pi/180
    } else if (event$char == 'Right') {
      angle <<- angle - 8 * pi/180
    } else if (event$char == 'Up') {
      # TODO check if already at topspeed
      at_max_speed <- FALSE
      if (!at_max_speed) {
        xvel <<- xvel + sin(-angle) / 1000
        yvel <<- yvel + cos(-angle) / 1000
      }
    }
  }

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Ship - Update position of the ship and then wrap in toroidal space
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  xpos <<- xpos + xvel
  ypos <<- ypos + yvel
  xpos <<- ifelse(xpos < 0, 1 + xpos, ifelse(xpos > 1, xpos - 1, xpos))
  ypos <<- ifelse(ypos < 0, 1 + ypos, ifelse(ypos > 1, ypos - 1, ypos))

  # Calc and draw new ship orientation + position
  ship_x <- ship_poly_x * cos(angle) - ship_poly_y * sin(angle) + unit(xpos, 'npc')
  ship_y <- ship_poly_y * cos(angle) + ship_poly_x * sin(angle) + unit(ypos, 'npc')
  grid::grid.polygon(
    x = ship_x,
    y = ship_y,
    gp = grid::gpar(fill = NA, col = 'white'),
  )

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Asteroids - update the position of the asteroids, then wrap in toroidal space
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  asxpos <<- asxpos + asxvel
  asypos <<- asypos + asyvel
  asxpos <<- ifelse(asxpos < 0, 1 + asxpos, ifelse(asxpos > 1, asxpos - 1, asxpos))
  asypos <<- ifelse(asypos < 0, 1 + asxpos, ifelse(asypos > 1, asypos - 1, asypos))

  # Draw asteroids
  for (i in seq(Nasteroids)) {
    ast_x <- asteroid_x + unit(asxpos[i], 'npc')
    ast_y <- asteroid_y + unit(asypos[i], 'npc')
    grid::grid.polygon(
      x = ast_x,
      y = ast_y,
      gp = grid::gpar(fill = NA, col = 'white'),
    )
  }
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Start the event loop. Press ESC to quit
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
run_loop(asteroids)
```


Since an interactive window cannot be captured in a vignette, a video
screen capture has been taken of the window and included below.

<video controls>
  <source src="images/asteroids.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 


