---
title: "Wordle"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Wordle}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(1)
```

```{r setup}
library(grid)
library(eventloop)
```

## Introduction

**Wordle** is a word guessing game that you can [play online](https://www.nytimes.com/games/wordle/index.html).

The goal is to guess a hidden target word by making guesses and 
receiving feedback on which letters in your guess are part of the target word.


# This Worldle game

This is a very 'lite' version of a Wordle game.

* No dictionary of words - any 5 letters in a row are considered a valid guess
* Hardcoded target word
* User must press "ENTER" to confirm word
* BACKSPACE is available


You could easily add all these things with a little bit of work, but I'm 
keeping this example simple so that the event logic is as clear as can be

## Setup

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# The colours
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
col_grey   <- grey(0.7)
col_green  <- "green"
col_yellow <- "yellow"


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Get a sequence of 5 colours given a guess word and a target word
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
get_colours <- function(guess, target) {
  
  # if the word is not yet complete, the return a vector of NAs
  # which will show up as 'blank' when filling the squares
  if (length(guess) == 0 || is.na(guess) || trimws(guess) == '' || nchar(guess) < 5) {
    return(rep(NA, 5))
  }
  
  guess  <- toupper(guess)
  target <- toupper(target)
  
  # Split the guess and target into individual letters
  this_cols <- rep(col_grey, 5)
  guess_letters <- strsplit(guess, '')[[1]]
  ref <- strsplit(target, '')[[1]]
  
  # Do the logic to get greens (easy)
  matched <- which(guess_letters == ref)
  this_cols[matched] <- col_green
  
  # Do the logic to figure out where the yellows are
  # This is more difficult
  ref[matched] <- '*'
  check_pos <- setdiff(1:5, matched)
  
  for (pos in check_pos) {
    letter <- guess_letters[pos]
    if (letter %in% ref) {
      this_cols[pos] <- col_yellow
      remove <- which(ref == letter)[1]
      ref[remove] <- '*'
    }
  }
  
  
  return(this_cols)
}



#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Positions for letters/squares
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
pos <- expand.grid(
  x = seq.int(0.2, 0.8, length.out = 5),
  y = seq.int(0.1, 0.9, length.out = 6)
)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Draw a board given guesses and target word
# No sanity checking is done here for number of guesses, length of target etc
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
draw_board <- function(guesses, target, current_wordnum) {
  lets <- unlist(strsplit(guesses, ''))
  length(lets) <- 30
  lets[is.na(lets)] <- ''
 
  cols <- unlist(lapply(guesses, get_colours, target = target))
  
  # ensure we don't highlight the current word while working on it.
  cols[5 * (current_wordnum - 1) + 1:5] <- NA_character_
  
  length(cols) <- 30
   
  grid.newpage()
  width <- height <- 0.13
  grid.rect(
    x      = pos$x, 
    y      = 1 - pos$y, 
    width  = width, 
    height = height, 
    default.units = 'snpc', 
    gp = gpar(fill = cols)
  )

  grid.text(lets, x = pos$x, y = 1 - pos$y, default.units = 'snpc',
            gp = gpar(cex = 2))
}
```


## Running the Game in an event loop

```{r eval = FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Global state for the eventloop
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
init    <- FALSE    # Has the puzzle been started?
solved  <- FALSE    # Has the puzzle been solved?
failed  <- FALSE    # Has the user completely failed to guess?
guesses <- c('')    # The guesses from the user
target  <- 'HELLO'  # The target word

current <- c() # the letters in the current guess
wordnum <- 1   # which attempt?


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' React to the user entering words and show win/loss status.
#' 
#' Press ESC to quit.
#' 
#' @param event The event from the graphics device. Is NULL when no event
#'        occurred.  Otherwise has `type` element set to:
#'        `event$type = 'mouse_down'` 
#'               - an event in which a mouse button was pressed
#'               - `event$button` gives the index of the button
#'        `event$type = 'mouse_up'`   
#'               - a mouse button was released
#'        `event$type = 'key_press'`  
#'               - a key was pressed
#'               - `event$char` holds the character as string
#'               - `event$int` holds the integer representation
#' @param mouse_x,mouse_y current location of mouse within window. If mouse is 
#'        not within window, this will be set to the last available coordinates
#' @param frame_num integer count of which frame this is
#' @param fps_actual,fps_target the curent framerate and the framerate specified
#'        by the user
#' @param dev_width,dev_height the width and height of the output device. Note:
#'        this does not cope well if you resize the window
#' @param ... any extra arguments ignored
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
wordle <- function(event, ...) {

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # If this is  first frame then the board *MUST* be drawn
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  
  if (!init) {
    init <<- TRUE
    draw_board(guesses, target, wordnum)
  }
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  
  # Build the current word a letter at a time
  # When the current word is 5 letters long, then add it to the list of guesses
  # Watch for the latest word matching target word => puzzle solved!
  # Watch for guesses exceeding 6 => user has failed!
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  
  if (!failed && !solved && !is.null(event)) {
    if (event$type == 'key_press') {
      if (event$char %in% c(letters, LETTERS) && length(current) < 5) {
        current <<- c(current, toupper(event$char))
        guesses[wordnum] <<- paste(current, collapse = '')
        draw_board(guesses, target, wordnum)
      } else if (event$int == 13 && length(current) == 5) { 
        # RETURN
        current <<- c()
        if (guesses[wordnum] == target) {
          solved <<- TRUE
          wordnum <<- wordnum + 1
          cat("Solved")
        } else {
          wordnum <<- wordnum + 1
          failed  <<- wordnum > 6
        }
        draw_board(guesses, target, wordnum)
      } else if (event$int == 8 && length(current) > 0) {
        # DELETE
        current <<- current[-length(current)]
        guesses[wordnum] <<- paste(current, collapse = '')
        draw_board(guesses, target, wordnum)
      }
    }
    if (solved) {
      grid.text(label = "Solved!", gp = gpar(col = 'blue', cex = 5))
    }
    if (failed) {
      grid.text(label = "Failed!", gp = gpar(col = 'red', cex = 5))
    }
  }

  
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Since we are only drawing in response to user input, we can do without 
# a double buffered display
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
eventloop::run_loop(wordle, double_buffer = FALSE)
```


Since an interactive window cannot be captured in a vignette, a video
screen capture has been taken of the window and included below.

<video controls>
  <source src="images/wordle.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video>


