---
title: "Mystery Curves"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mystery Curves}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Interactive plotting of mystery curves

George Savva outlined a great creative coding technique he uses in his [post on Mystery Curves](https://georgemsavva.github.io/creativecoding/posts/mystery/).

These curves are the superposition of multiple circular paths and render as pleasing 
spiral/repeating patterns.

In this example the mouse X and Y position are used to control two key 
parameters in the rendering - making this an interactive exploration of the parameter space.


## Controls

* Mouse position is used to control two key curve parameters
* Press ESC to quit


```{r setup, eval=FALSE}
library(grid)
library(eventloop)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' The main 'draw' function called within the eventloop
#'
#' @param event The event from the graphics device. Is NULL when no event
#'        occurred.  Otherwise has `type` element set to:
#'        `event$type = 'mouse_down'` 
#'               - an event in which a mouse button was pressed
#'               - `event$button` gives the index of the button
#'        `event$type = 'mouse_up'`   
#'               - a mouse button was released
#'        `event$type = 'mouse_move'`   
#'               - mouse was moved 
#'        `event$type = 'key_press'`  
#'               - a key was pressed
#'               - `event$str` String describing which key was pressed. 
#'                  See \code{grDevices::setGraphicsEventHandlers} for more information.
#' @param mouse_x,mouse_y current location of mouse within window in normalised 
#'        coordinates in the range [0, 1]. If mouse is 
#'        not within window, this will be set to the last available coordinates
#' @param frame_num Current frame number (integer)
#' @param fps_actual,fps_target the curent framerate and the framerate specified
#'        by the user
#' @param dev_width,dev_height the width and height of the output device. Note:
#'        this does not cope well if you resize the window
#' @param ... any extra arguments ignored
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
circle <- function(amp, freq, phase) amp*1i^(freq*seq(0,600,l=260)+phase)

draw <- function(event, mouse_x, mouse_y, ...) {
  
  grid.rect(gp = gpar(fill = '#04010F'))

  a <- mouse_x * 5;
  l <- sin(pi*(2*a-.5))+1

  z <-circle(mouse_y, 1, 0) +
    circle(l, ceiling(a), -8*a) +
    circle(l/2-1,ceiling(((-a+2.5)%%5)-5), -4*a)

  x <- Re(z)/7 + 0.5
  y <- Im(z)/7 + 0.5

  hue <- (a+(Re(z/10))) %% 1
  col <- hsv(hue, 0.65, 1, .1)

  grid.points(x, y, gp = gpar(col = col), pch = 20, default.units = 'npc')

  grid.segments(x, y, c(x[-1], x[1]), c(y[-1], y[1]), gp = gpar(col = col, lwd = 2))
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Start the event loop. Press ESC to quit
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
eventloop::run_loop(draw, fps_target = 20, double_buffer = TRUE, show_fps = TRUE)
```

Since an interactive window cannot be captured in a vignette, a video
screen capture has been taken of the window and included below.

<video controls>
  <source src="images/mystery-curves.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 

