---
title: "Realtime Plot Updates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Realtime Plot Updates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Realtime stream-based plotting

Plot new lines at every timestep as if drawing data from a sensor


## Controls

* The behaviour of this callback function is not influenced by anything the
  user does.
* Press ESC to quit


```{r eval = FALSE}
library(grid)
library(eventloop)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Initialise a number of data streams
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
N        <- 20
Nstreams <- 100
streams  <- lapply(seq(Nstreams), function(x) rep(0.5, N))
alpha    <- 0.05
cols     <- rainbow(Nstreams)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' The plotting/behaviour at each call doesn't actually depend on anything
#' the user is doing, so all arguments are ignored
#'
#' @param event The event from the graphics device. Is NULL when no event
#'        occurred.  Otherwise has `type` element set to:
#'        `event$type = 'mouse_down'` 
#'               - an event in which a mouse button was pressed
#'               - `event$button` gives the index of the button
#'        `event$type = 'mouse_up'`   
#'               - a mouse button was released
#'        `event$type = 'mouse_move'`   
#'               - mouse was moved 
#'        `event$type = 'key_press'`  
#'               - a key was pressed
#'               - `event$char` holds the character as string
#'               - `event$int` holds the integer representation
#' @param mouse_x,mouse_y current location of mouse within window. If mouse is 
#'        not within window, this will be set to the last available coordinates
#' @param frame_num integer count of which frame this is
#' @param fps_actual,fps_target the curent framerate and the framerate specified
#'        by the user
#' @param dev_width,dev_height the width and height of the output device. Note:
#'        this does not cope well if you resize the window
#' @param ... any extra arguments ignored
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
monitor_status <- function(...) {
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Update all streams by pulling a new value from the source.
  # In this simple example, the datasource is just 'runif(1)'
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  streams <<- lapply(
    streams,
    function(stream) {
      c(
        stream[2:N],
        alpha * runif(1) + (1 - alpha) * stream[N]
      )
    }
  )

  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Clear the screen
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  grid.rect(gp = gpar(fill = 'white'))
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Plot the data streams in different colours
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  for (i in seq_along(streams)) {
    grid.lines(
      x = seq.int(0, 1, length.out = N),
      y = streams[[i]],
      gp = gpar(col = cols[[i]])
    )
  }
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Start the event loop. Press ESC to quit
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
run_loop(monitor_status, show_fps = TRUE, fps_target = 30)
```


Since an interactive window cannot be captured in a vignette, a video
screen capture has been taken of the window and included below.

<video controls>
  <source src="images/plotting.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 


