---
title: "Particles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Particles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This example shows the simulated movement of particles.  This is not really
anything physics-based or reality-based, but looks a little bit organic - like
single celled creatures roaming around.

There is no interactivity in this example - but could be interesting to add e.g.
adjust velocity vectors away from the current mouse position to make it look 
like the particles are avoiding it.


```{r setup}
library(grid)
library(eventloop)
```



```{r eval=FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Particle parameters
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
N <- 100
x  <- runif(N)
y  <- runif(N)
vx <- runif(N, -0.5, 0.5) / 160
vy <- runif(N, -0.5, 0.5) / 160


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Update factor
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
k <- 0.01


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Main event callback
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
update_particles <- function(...) {
  
  # Mean velocity of a random half of the population
  mean_vx <- mean(sample(vx, N/2))
  mean_vy <- mean(sample(vy, N/2))
  
  # Update the velocity towards the mean + noise
  vx <<- (1 - k) * vx + k * mean_vx + runif(N, -0.5, 0.5) / 400
  vy <<- (1 - k) * vy + k * mean_vy + runif(N, -0.5, 0.5) / 400
  
  # Update position
  x <<- x + vx
  y <<- y + vy
  
  # Wrap around coordinates
  x[x < 0] <<- 1
  x[x > 1] <<- 0
  y[y < 0] <<- 1
  y[y > 1] <<- 0
  
  # Overlay a white rect with alpha to get a particle trail effect
  grid.rect(gp = gpar(fill = '#ffffff80'))
  grid.points(x, y, default.units = 'npc', gp = gpar(fill = 'grey80'), pch = 21)
}


eventloop::run_loop(update_particles, show_fps = TRUE, width = 10, height =7)
```




Since an interactive window cannot be captured in a vignette, a video
screen capture has been taken of the window and included below.

<video controls>
  <source src="images/particles.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 


