---
title: "Drawing with Lines"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Drawing with Lines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## A pixel-based drawing canvas

Draw lines as the mouse moves.

Extra logic is used to store prior coordinates so 
  that the output is a continuous line when drawing


## Controls

* Press and hold the left mouse button to draw
* Press SPACE to clear the canvas
* Press ESC to quit


```{r setup, eval=FALSE}
library(grid)
library(eventloop)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Set up the global variables which store the state of the world
#  'drawing'  Currently drawing?
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
drawing <- FALSE
last_x  <- NA
last_y  <- NA

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' The main 'draw' function called within the eventloop
#'
#' @param event The event from the graphics device. Is NULL when no event
#'        occurred.  Otherwise has `type` element set to:
#'        `event$type = 'mouse_down'` 
#'               - an event in which a mouse button was pressed
#'               - `event$button` gives the index of the button
#'        `event$type = 'mouse_up'`   
#'               - a mouse button was released
#'        `event$type = 'mouse_move'`   
#'               - mouse was moved 
#'        `event$type = 'key_press'`  
#'               - a key was pressed
#'               - `event$str` String describing which key was pressed. 
#'                  See \code{grDevices::setGraphicsEventHandlers} for more information.
#' @param mouse_x,mouse_y current location of mouse within window in normalised 
#'        coordinates in the range [0, 1]. If mouse is 
#'        not within window, this will be set to the last available coordinates
#' @param frame_num Current frame number (integer)
#' @param fps_actual,fps_target the curent framerate and the framerate specified
#'        by the user
#' @param dev_width,dev_height the width and height of the output device. Note:
#'        this does not cope well if you resize the window
#' @param ... any extra arguments ignored
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
draw <- function(event, mouse_x, mouse_y, ...) {
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Process events
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (!is.null(event)) {
    if (event$type == 'mouse_down') {
      drawing <<- TRUE
    } else if (event$type == 'mouse_up') {
      drawing <<- FALSE
      last_x <<- NA
      last_y <<- NA
    } else if (event$type == 'key_press' && event$str == ' ') {
      grid::grid.rect(gp = gpar(col=NA, fill='white')) # clear screen
      last_x <<- NA
      last_y <<- NA
    }
  }
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # If drawing mode is active, draw from the last point to this point.
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (drawing) {
    if (!is.na(last_x)){
      grid::grid.lines(
        x = c(last_x, mouse_x),
        y = c(last_y, mouse_y)
      )
    }
    
    last_x <<- mouse_x
    last_y <<- mouse_y
  }
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Start the event loop. Press ESC to quit
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
eventloop::run_loop(draw, fps_target = NA, double_buffer = FALSE)
```

Since an interactive window cannot be captured in a vignette, a video
screen capture has been taken of the window and included below.

<video controls>
  <source src="images/basic-canvas-smooth.mp4" type="video/mp4">
  Your browser does not support the video tag.
</video> 

