---
title: "Verbose Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Verbose Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Verbose Example

The following code is a verbose eventloop callback which just prints the 
information about the current event to the console.

This could be a good place to start if developing code from scratch.


```{r eval = FALSE}
library(eventloop)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Global variables used to retain information between calls
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
last_x <- NA
last_y <- NA


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' This is a verbose callback that prints information to the console
#' about the latest event
#' 
#' Press ESC to quit.
#' 
#' @param event The event from the graphics device. Is NULL when no event
#'        occurred.  Otherwise has `type` element set to:
#'        `event$type = 'mouse_down'` 
#'               - an event in which a mouse button was pressed
#'               - `event$button` gives the index of the button
#'        `event$type = 'mouse_up'`   
#'               - a mouse button was released
#'        `event$type = 'mouse_move'`   
#'               - mouse was moved 
#'        `event$type = 'key_press'`  
#'               - a key was pressed
#'               - `event$str` String describing which key was pressed. 
#'                  See \code{grDevices::setGraphicsEventHandlers} for more information.
#' @param mouse_x,mouse_y current location of mouse within window in normalised 
#'        coordinates in the range [0, 1]. If mouse is 
#'        not within window, this will be set to the last available coordinates
#' @param frame_num Current frame number (integer)
#' @param fps_actual,fps_target the curent framerate and the framerate specified
#'        by the user
#' @param dev_width,dev_height the width and height of the output device. Note:
#'        this does not cope well if you resize the window
#' @param ... any extra arguments ignored
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
verbose <- function(event, mouse_x, mouse_y, frame_num, fps_actual,
                    fps_target, dev_width, dev_height, ...) {
  
  # Check if an mouse or keyboard event happened.
  if (!is.null(event)) {
    
    # Draw a line from the last mouse position to the current position.
    if (!is.na(last_x)) {
      grid::grid.lines(
        x = c(last_x, mouse_x),
        y = c(last_y, mouse_y)
      )
    }
    
    # Keep track of latest mouse position in the global variables
    last_x <<- mouse_x
    last_y <<- mouse_y
    
    # "Handle" the event by printing to console
    co <- sprintf("Screen [%i, %i] Mouse Loc: [%.2f, %.2f]", dev_width, dev_height, mouse_x, mouse_y)
    if (event$type %in% c('mouse_down', 'mouse_up', 'mouse_move')) {
      cat(co, "Mouse: ", event$type, " - button: [", event$button, "]\n")
    } else if (event$type == 'mouse_move') {
      cat(co, "Move \n")
    } else if (event$type == 'key_press') {
      cat(co, "Key: ", event$str, "\n")
    }
  }
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Start the event loop. Press ESC to quit.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
eventloop::run_loop(verbose, fps_target = 10, show_fps = TRUE)
```



## Verbose Reference Example - R6

It is not always wide to put lots of information in the global environment, 
so another way of storing information is to use an environment.

`{R6}` objects are a user-friendly environment-based way of storing functions 
and values.  In the following example, an R6 method is used as the
user callback function to the eventloop.


```{r setup, eval = FALSE}
library(R6)
library(eventloop)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# R6 environment object
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
Agent <- R6::R6Class(
  "Agent",
  
  public = list(
    
    last_x = NULL,
    last_y = NULL,
    
    initialize = function() {
      self$last_x <- NA
      self$last_y <- NA
      invisible(self)
    },
    
    update = function(event, mouse_x, mouse_y, frame_num, fps_actual,
                      fps_target, dev_width, dev_height, ...) {
      
      # Did an event happen at all in this window?
      if (!is.null(event)) {
        
        # Draw a line
        if (!is.na(self$last_x)) {
          grid::grid.lines(
            x = c(self$last_x, mouse_x),
            y = c(self$last_y, mouse_y)
          )
        }
        
        # Keep track of latest mouse position
        self$last_x <- mouse_x
        self$last_y <- mouse_y
        
        # "Handle" the event by printing to console
    co <- sprintf("Screen [%i, %i] Mouse Loc: [%.2f, %.2f]", dev_width, dev_height, mouse_x, mouse_y)
        if (event$type %in% c('mouse_down', 'mouse_up')) {
          cat(co, "Mouse: ", event$type, " - button: [", event$button, "]\n")
        } else if (event$type == 'mouse_move') {
          cat(co, "Move \n")
        } else if (event$type == 'key_press') {
          cat(co, "Key: ", event$str, "\n")
        }
      }
    }
  )
)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create the 'Agent' object and start the event loop. Press ESC to quit.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
agent <- Agent$new()
eventloop::run_loop(agent$update, fps_target = 10, show_fps = TRUE)
